This work is not mine alone. Several people have contributed to the work that
went into this document, either directly or indirectly.
Firstly, I am immensely grateful to my advisor, Eduard Hovy, for teaching me
to never shy away from, and even get excited about working on important, yet frustratingly hard problems. While I admit that the problems that ended up being solved
in this thesis are only a fraction of the ones that I had set out to solve, that I even
attempted to tackle them is largely because of him. I am also thankful to Chris Dyer,
William Cohen, and Luke Zettlemoyer. I could not have asked for a better committee. I was fortunate to have received valuable insights from them in shaping my
thesis, limited only by my ability to utilize them.

I learned a great deal from my talented collaborators and mentors: Waleed Ammar,
Matt Gardner, and Jayant Krishnamurthy. Waleed stood by me through several failed
attempts at getting the ontology-aware encoders to work, and Jayant and Matt are
largely responsible for driving me towards semantic parsing. For all this, I am greatly
thankful to them. I spent a large chunk of my time as a PhD student visiting the Allen
Institute for Artificial Intelligence, and I am highly grateful to Oren Etzioni and the
rest of AI2, both for letting me stay, and for inspiring me to work on impactful
problems.

My friends have made my time over the last five years much more enjoyable.
Thanks to Sujay, Chandan, Swabha, and Suchin for the fun board game and movie
nights, brunches, and hikes, both in Pittsburgh and in Seattle. Thanks to Kartik,
Keerthi, and Mrinmaya for the numerous lunches and workout sessions.

I am thankful to my parents for teaching me the value of knowledge, instilling
in me the right priorities, and making me a person capable of getting a PhD. I would
also like to thank my brother, Praveen, for driving me towards research in Computer
Science, and for being a voice of sanity and a source of wisdom. Finally, I would
like to thank my dear wife Modhurima, for providing me a solid emotional grounding
over the past four years, for giving meaning to life outside of work, and for all the
love.
